Designing Qusecs Optimally: Quasi-uniform or Self-similar Layered Materials as Recursively Decomposed Solutions of Geometric Constraint Systems

### 2014-12-20 doc/GMP2015/appendix_a.tex
Changed proof of algo. Corrected references. Increased references in intro. Changed complexity of algo throughout paper. Added citation in intro. Other smaller changes.

\begin{proof}
Remark \ref{lemma:union_intersection} shows that the intersection of any two isostatic subgraphs (such as $C_i$ and $C_j$) can only result in trivial or isostatic subgraphs. Therefore, these are the only possibilities to consider.

@@

Take the strict subset $S\subsetneq \{1,\ldots,N\}$ such that $U=\bigcup_{i\in S}{C_i}$ which we \vemph{assume} is isostatic. If $U\neq C$, then we just found a larger proper subgraph and the children were not vertex-maximal to begin with. So, it must be that $U=C$.
\usestwod
However, since $C_i \cap C_j$ is trivial then for $k\notin S$ we know, by Lemma \ref{lemma:uc_intersection_makes_all_uc}, $U\cap C_k$ must be one or more trivial subgraphs. By definition of a DR-plan $C_k=C\cap C_k$ and we know that $U=C$ so $C_k=U\cap C_k$. Thus, $C_k$ is both one or more trivial subgraphs and a isostatic subgraph of $C$. Thus, the assumption is wrong and $U$ cannot be isostatic. As $C$ is isostatic, this means no proper subset of the children in ComDRP$(C)$ can union to form $C$.

\medskip\noindent
\vemph{Part 2:} If $C_i \cap C_j$ is isostatic, then, by Remark \ref{lemma:union_intersection}, $C_i \cup C_j$ is also isostatic. This means that, by Lemma \ref{lemma:wc_intersection_makes_all_wc}, any two children of $C$ will union to $C$ itself. Thus, any two children are potential choices for the optimal DR-plan as they all create equal fan-in (exactly two) at this level.

@@

\begin{proof}
Lemma \ref{lemma:wc_intersection_is_C} shows that $C_i\cup C_j$ must equal the entire parent $C$.
\end{proof}

@@

The alternative phrasing is due to Lemma \ref{lemma:wc_intersection_is_C}.

@@

    \item 1 case: $C_k=\inducedOnC{R'_i\cup R'_j\cup D_{i,j}}$ is not possible. Since $C_i\cup C_k = \inducedOnC{R'_i\cup R_j\cup D_{i,j}}\neq C$ we have from Lemma \ref{lemma:wc_intersection_is_C} that $C_i\cup C_k$ cannot be isostatic. We also know it cannot be trivial because it contains isostatic subgraphs. This means it must be under-constrained. From Lemma \ref{lemma:union_intersection}, we know that $C_i\cap C_k=\inducedOnC{R'_j\cup D_{i,j}}$ must then be trivial. This is impossible because $D_{i,j}$ is isostatic, thereby contradicting the assumption that $C_k$ is isostatic.

@@

\begin{proof}
Assume there is some $k$ such that $C_i\cap C_k$ is not trivial. By Remark \ref{lemma:union_intersection}, $C_i\cap C_k$ must be isostatic. Then, by Lemma \ref{lemma:wc_intersection_makes_all_wc}, the intersection between any two children must be isostatic. This means that $C_i\cap C_j$ is isostatic. Therefore, such a $k$ cannot exist and all intersections are trivial.
\end{proof}

@@

\begin{proof}
The first step of the algorithm is finding the isostatic vertex-maximal proper subgraphs of the input isostatic graph $G$. Do this by first dropping arbitrary edge $e$ from the edge set and running the pebble game algorithm \cite{Jacobs:1997:PG} on this subgraph, which is $O(|V|^2)$. The output of this will be the list of component-candidates. Then run the Frontier algorithm \cite{Oung:2001:FFE:376957.376995} \todo{complexity?} on each candidate with edge $e$, which will find the minimal subgraph containing the candidate and $e$, which we call $D$. If $D=G$, move $D$ to the true-component list and remove the candidate from the candidate-component list. If $D\subsetneq G$, remove all candidates that are subgraphs of $D$ from the candidate-component list. Then, check the union of $D$ with the all items in the true-component list; if the union is isostatic, halt and output just these two components as children. Otherwise, run to completion and output all items in the true-component list.

This is run recursively on each node. If you use dynamic programming to store the decompositions of each component, it prevents repetition of work. In this case you will have to do work on no more than $O(V)$ nodes that get smaller at each level.

Thus, the running time is at most $O(|V|^3)$. \todo{does this amortize correctly?}
\end{proof}

@@

\subsection{Proofs from Section \ref{sec:pinnedline}}
\subsubsection{Proof of Lemma \ref{lemma:wc_intersection_makes_all_wc} --- For 2-dimensional pinned line incidence graphs}
\label{sec:appendix_pinned}

\begin{proof} %[Proof of Lemma~\ref{lemma:wc_intersection_makes_all_wc} for 2D pinned line incidence graphs]
We use the same notations as in the original proof of Lemma~\ref{lemma:wc_intersection_makes_all_wc} in the last section.

Then we analyze all the possible cases for $C_k$.
\begin{itemize}
    \item 1 case: $C_k=Idc(C,R'_i\cup R'_j\cup D_{i,j})$ is not possible. Since $C_k\cup C_i = Idc(C,R'_i\cup R_j\cup D_{i,j})\neq C$ we have from Lemma \ref{lemma:wc_intersection_is_C} that $C_k\cup C_i$ cannot be isostatic.
    By Lemma~\ref{lem:pinned_union_intersection}, it must be under-constrained,
    so one of $C_i$ and $C_k$ must be an edge, contradicting the assumption that both $C_i$ and $C_k$ are isostatic.

@@

\subsubsection{Proof of Remark \ref{rem:bodypin_is_barjoint}}
\begin{proof}
    We can replace each body that has only one pin by a single vertex. A body with 2 pins can be replaced by an edge. In general, a body with $n$ pins can be replaced by a 2-tree on $n$ vertices. When looking for a DR-plan, we treat each body as trivial, so they become the leaves of our plan. The optimal completion problem and approach of Section \ref{sec:DRP} are unchanged. The optimal parameterization problem in Section \ref{sec:DRP} now has an additional constraint that all edges in the 2-tree representation of the bodies must be removed together, not individually.
\end{proof}

\subsubsection{Proof of Observation \ref{obs:bodypin_drp}}
\begin{proof}
    \todo{Item 1 Follows from paper?}

    For item 2, consider the case where we have more than 2 proper vertex maximal 1-dof sub-graphs $s_1, ..., s_k, k > 2$. Then, if $k_i$ and $k_j$ are joined by $2$ pins, $k_i \cup k_j$ would be $(3,4)$-tight and hence $k_i$ and $k_j$ are not vertex maximal.
\end{proof}

\subsubsection{Proof of Observation \ref{thm:1dofcase}}
\begin{proof}
    Suppose we are given a body-pin graph and its corresponding body-bar graph $G$ and have obtained the 1-dof DR-plan $T$. Each node of $T$ will then be a vertex maximal proper 1-dof sub-graph of $G$.

@@

    Thus, say we have a $T$ as described in the proof for Theorem \ref{thm:1dofcase}. When we pin a node $b$, we will have the same structure as before. Suppose we are looking at an isostatic node $v$ after pinning $b$. Then, the children of $v$ (except one if $v \neq b$) will be 1-dof. The complexity for that node is then just that of solving each of its children. In general then, the number of parameters for that node will be $np(v) = 4nc_1(v)+3$, if $v \neq b$ and $np(b) = 4nc_1(b)$, where $nc_k(v)$ is the number of $k$-dof children of $v$.

@@

>>>>

\begin{proof}
Observation \ref{lemma:union_intersection} shows that the intersection of any two isostatic subgraphs (such as $C_i$ and $C_j$) can only result in trivial or isostatic subgraphs. Therefore, these are the only possibilities to consider.

@@

Take the strict subset $S\subsetneq \{1,\ldots,N\}$ such that $U=\bigcup_{i\in S}{C_i}$ which we \vemph{assume} is isostatic. If $U\neq C$, then we just found a larger proper subgraph and the children were not vertex-maximal to begin with. So, it must be that $U=C$.
\usestwod
However, since $C_i \cap C_j$ is trivial then for $k\notin S$ we know, by Lemma \ref{lemma:combined_lemma}, point \ref{lemma:uc_intersection_makes_all_uc}, $U\cap C_k$ must be one or more trivial subgraphs. By definition of a DR-plan $C_k=C\cap C_k$ and we know that $U=C$ so $C_k=U\cap C_k$. Thus, $C_k$ is both one or more trivial subgraphs and a isostatic subgraph of $C$. Thus, the assumption is wrong and $U$ cannot be isostatic. As $C$ is isostatic, this means no proper subset of the children in ComDRP$(C)$ can union to form $C$.

\medskip\noindent
\vemph{Part 2:} If $C_i \cap C_j$ is isostatic, then, by Observation \ref{lemma:union_intersection}, $C_i \cup C_j$ is also isostatic. This means that, by Lemma \ref{lemma:combined_lemma}, point \ref{lemma:wc_intersection_makes_all_wc}, any two children of $C$ will union to $C$ itself. Thus, any two children are potential choices for the optimal DR-plan as they all create equal fan-in (exactly two) at this level.

@@

\begin{proof}
Lemma \ref{lemma:combined_lemma}, point \ref{lemma:wc_intersection_is_C}, shows that $C_i\cup C_j$ must equal the entire parent $C$.
\end{proof}

@@

The alternative phrasing is due to Lemma \ref{lemma:combined_lemma}, point \ref{lemma:wc_intersection_is_C}.

@@

    \item 1 case: $C_k=\inducedOnC{R'_i\cup R'_j\cup D_{i,j}}$ is not possible. Since $C_i\cup C_k = \inducedOnC{R'_i\cup R_j\cup D_{i,j}}\neq C$ we have from Lemma \ref{lemma:combined_lemma}, point \ref{lemma:wc_intersection_is_C}, that $C_i\cup C_k$ cannot be isostatic. We also know it cannot be trivial because it contains isostatic subgraphs. This means it must be under-constrained. From Observation \ref{lemma:union_intersection}, we know that $C_i\cap C_k=\inducedOnC{R'_j\cup D_{i,j}}$ must then be trivial. This is impossible because $D_{i,j}$ is isostatic, thereby contradicting the assumption that $C_k$ is isostatic.

@@

\begin{proof}
Assume there is some $k$ such that $C_i\cap C_k$ is not trivial. By Observation \ref{lemma:union_intersection}, $C_i\cap C_k$ must be isostatic. Then, by Lemma \ref{lemma:combined_lemma}, point \ref{lemma:wc_intersection_makes_all_wc}, the intersection between any two children must be isostatic. This means that $C_i\cap C_j$ is isostatic. Therefore, such a $k$ cannot exist and all intersections are trivial.
\end{proof}

@@

\begin{proof}
The first step of the algorithm is finding the isostatic vertex-maximal proper subgraphs of the input isostatic graph $G$. Do this by first dropping arbitrary edge $e$ from the edge set and running the pebble game algorithm \cite{Jacobs:1997:PG} on this subgraph, which is $O(|V|^2)$. The output of this will be a list of component-candidates. Then run the Frontier algorithm \cite{hoffman2001decompositionII} \cite{lomonosov2004graph} ($O(|V|)$) on each candidate along with edge $e$, which will find the minimal subgraph $D$ that contains the candidate and $e$. If $D=G$, move the candidate (before adding $e$) to the true-component list and remove it from the candidate-component list. If $D\subsetneq G$, remove all candidates that are subgraphs of $D$ from the candidate-component list and add $D$ back to the list. Continue with the next candidate. When you exhaust the candidates, move to the next step.

Test the intersection of any two of the true-components. If the intersection is trivial, the entire list becomes the children. Recursively apply the algorithm to each child.

If the intersection is non-trivial (necessarily isostatic), immediately compute the sequential decomposition down to $I=\bigcap_{k\in S_N}{C_k}$. The root of this idea is discussed in detail in the proof of Theorem \ref{theorem:main}, and we use the notation here. We construct something that, while recombining it, results in the same systems to solve as the true DR-plan. Begin with the parent $C$. Its children will be $C\setminus R_1$ and $R_1$ plus the nodes and edges adjacent to $R_1$ that lie in $I$. Recursively apply the algorithm to the child containing $R_1$. The children of $C\setminus R_1$ will be $C\setminus (R_1\cup R_2)$ and $R_2$ plus the nodes and edges adjacent to $R_2$ that lie in $I$. Etc. At $N$ levels down, the 2 children will be $I$ and $R_N$ plus the nodes and edges adjacent to $R_N$ that lie in $I$. Apply the algorithm recursively to both children.
% Although the nodes containing $R_1,\ldots,R_N$ are under-constrained, simply take the children to be the roots of the forest

The reason we construct it this way is because we end up with a tree instead of a directed acyclic graph. The leaves of this tree will be exactly every single edge in the graph, with none repeated. Therefore, the number of nodes in the tree is $O(|E|)$ which, for isostatic input graphs, is $O(|V|)$.

Therefore, the complexity of the algorithm as a whole is $O(|V|^3)$.
%
% \note In the best case, this can be $O(V^2)$ if the graph is a 2-tree
\end{proof}

% \begin{proof}
% The first step of the algorithm is finding the isostatic vertex-maximal proper subgraphs of the input isostatic graph $G$. Do this by first dropping arbitrary edge $e$ from the edge set and running the pebble game algorithm \cite{Jacobs:1997:PG} on this subgraph, which is $O(|V|^2)$. The output of this will be a list of component-candidates. Then run the Frontier algorithm \cite{hoffman2001decompositionII} \cite{lomonosov2004graph} ($O(|V|)$) on each candidate along with edge $e$, which will find the minimal subgraph $D$ that contains the candidate and $e$. If $D=G$, move $D$ to the true-component list and remove the candidate from the candidate-component list. If $D\subsetneq G$, check the union of $D$ with the all items in the true-component list; if the union is isostatic, halt and output just these two components as children. Otherwise, remove all candidates that are subgraphs of $D$ from the candidate-component list and add $D$. Continue with the next candidate. If you exhaust the candidates, output the entire true-component list.

% %on the children
% This is run recursively on each node. If you use dynamic programming to store the decompositions of each component, it prevents repetition of work (done by using the DAG representation of the DR-plan). In this case you will have to do work on no more than $O(|V|)$ nodes that get smaller at each level.

% Thus, the running time is at most $O(|V|^3)$. \todo{does this amortize correctly?}

% It's a tree in only trivial case -> O(V), always order of leaves
% If its isostatic intersection, we do the sequential extension,intersections only done once.

% inductively. The number of nodes at this level will e more than the total. Number of leaves (terminals for dag) is at most the number of edges

% \end{proof}

@@

@@

\subsubsection{Proof of Remark \ref{rem:bodypin_is_barjoint}}
\begin{proof}
    We can replace each body that has only one pin by a single vertex. A body with 2 pins can be replaced by an edge. In general, a body with $n$ pins can be replaced by a 2-tree on $n$ vertices. When finding a DR-plan, we treat each body as trivial, so they become the leaves of our plan. The optimal recombination problem and approach of Section \ref{sec:recomb} are unchanged. The optimal completion via the optimal modification problem in Section \ref{sec:recomb} now has an additional constraint that all edges in the 2-tree representation of the bodies must be removed together, not individually.
\end{proof}

\subsubsection{Proof of Observation \ref{obs:bodypin_drp}}
\begin{proof}
    The existence of this canonical DR-plan follows from the same arguments as in the proof of Theorem \ref{theorem:main}. The only difference is the definition of a trivial intersection. In this case, when two subgraphs share more than 1 body, they become rigid (in fact over constrained). Sharing a pin is not considered an intersection. Such a structure is viewed as two subgraphs each sharing 1 body with a 3rd 1-dof subgraph which essentially just consists of those two bodies pinned together.
\end{proof}

\subsubsection{Proof of Remark \ref{thm:1dofcase}}
\begin{proof}
    Suppose we are given a body-pin graph and its corresponding body-bar graph $G$ and have obtained the 1-dof DR-plan $T$. Each node of $T$ will then be a vertex maximal proper 1-dof sub-graph of $G$.

@@

    Thus, say we have a $T$ as described in the proof for Remark \ref{thm:1dofcase}. When we pin a node $b$, we will have the same structure as before. Suppose we are looking at an isostatic node $v$ after pinning $b$. Then, the children of $v$ (except one if $v \neq b$) will be 1-dof. The complexity for that node is then just that of solving each of its children. In general then, the number of parameters for that node will be $np(v) = 4nc_1(v)+3$, if $v \neq b$ and $np(b) = 4nc_1(b)$, where $nc_k(v)$ is the number of $k$-dof children of $v$.

@@

\subsection{Proofs from Section \ref{sec:pinnedline}}
\subsubsection{Proof of Lemma \ref{lemma:combined_lemma}, Point \ref{lemma:wc_intersection_makes_all_wc} --- For 2-dimensional pinned line incidence graphs}
\label{sec:appendix_pinned}

\begin{proof} %[Proof of Lemma~\ref{lemma:wc_intersection_makes_all_wc} for 2D pinned line incidence graphs]
We use the same notations as in the original proof of Lemma \ref{lemma:combined_lemma}, Point \ref{lemma:wc_intersection_makes_all_wc}, in the last section.

Then we analyze all the possible cases for $C_k$.
\begin{itemize}
    \item 1 case: $C_k=Idc(C,R'_i\cup R'_j\cup D_{i,j})$ is not possible. Since $C_k\cup C_i = Idc(C,R'_i\cup R_j\cup D_{i,j})\neq C$ we have from Lemma \ref{lemma:combined_lemma}, Point \ref{lemma:wc_intersection_is_C}, that $C_k\cup C_i$ cannot be isostatic.
    By Lemma~\ref{lem:pinned_union_intersection}, it must be under-constrained,
    so one of $C_i$ and $C_k$ must be an edge, contradicting the assumption that both $C_i$ and $C_k$ are isostatic.

### 2014-12-20 doc/GMP2015/s00_abstract.tex
Changed proof of algo. Corrected references. Increased references in intro. Changed complexity of algo throughout paper. Added citation in intro. Other smaller changes.

Crucial for the design and analysis of the abovementioned properties
is the recursive decomposition of diverse types of underlying 2D
geometric constraint systems that are \vemph{isostatic or
underconstrained (which we call collectively as \dfn{qusecs})}. While
the optimal recursive decomposition (\dfn{DR-planning}) problem is NP-
hard even for (possibly overconstrained) 2D bar-joint constraint
systems we describe an $O(n^2)$ algorithm  for qusecs. Besides
elucidating NP-hardness, the algorithm outputs \dfn{canonical} DR-
plans that meet various desirable, previously studied criteria. We
leverage recent results on Cayley configuration spaces to show that
the indecomposable systems -  that are solved at node of the optimal
DR-plan by recombining solutions to child systems - can be minimally
modified to become decomposable and have a small DR-plan, leading to
efficient realization algorithms. We show formal connections to well-
known problems such as completion of underconstrained systems. We
briefly discuss the modeling and analysis of two specific 2D layered
materials as applications of the above theory and algorithms. We
intend to make a software implementation and videos available upon
request and publicly available for the final version.
\end{abstract}

>>>>

Crucial for the design and analysis of the abovementioned properties
is the recursive decomposition of diverse types of underlying 2D
geometric constraint systems that are \vemph{isostatic or
underconstrained (which we call collectively as \dfn{qusecs})}. While
the optimal recursive decomposition (\dfn{DR-planning}) problem is NP-
hard even for (possibly overconstrained) 2D bar-joint constraint
systems we describe an \candrpcomplexity algorithm  for qusecs. Besides
elucidating NP-hardness, the algorithm outputs \dfn{canonical} DR-
plans that meet various desirable, previously studied criteria. We
leverage recent results on Cayley configuration spaces to show that
the indecomposable systems -  that are solved at node of the optimal
DR-plan by recombining solutions to child systems - can be minimally
modified to become decomposable and have a small DR-plan, leading to
efficient realization algorithms. We show formal connections to well-
known problems such as completion of underconstrained systems. We
briefly discuss the modeling and analysis of two specific 2D layered
materials as applications of the above theory and algorithms. We
intend to make a software implementation and videos available upon
request and publicly available for the final version.
\end{abstract}

### 2014-12-20 doc/GMP2015/s01_introduction.tex
Changed proof of algo. Corrected references. Increased references in intro. Changed complexity of algo throughout paper. Added citation in intro. Other smaller changes.

Examples of such materials (See Figure \ref{fig:material_examples})
include:

\begin{enumerate}
    \item Cross-sections of microtubule structures~\cite{microtubule_necklace} (Figure \ref{fig:material_examples:microtubule}) e.g., in ciliary membranes and transitions~\cite{microtubule_cilia}

    \item Cross-sections of organic tissue with hierarchical structure, e.g. compact bone and tendon (Figure \ref{fig:material_examples:tendon}) \uncited

    \item Crosslinked cellulose or collagen microfibril monolayers e.g., in cell-walls~\cite{wikimediacommons2010afm}~\cite{wikimediacommons2007plant}, as well as crosslinked actin filaments in the cytoskeleton matrix. See Section \ref{sec:pinnedline}.

    \item More recent, engineered examples, including disordered
    graphene layers~\cite{Graphene1}~\cite{Graphene2} sometimes
    reinforced by microfibrils; and DNA assemblies including a recent
    Sierpinski gasket~\cite{self_assembly_sierpinski}, bringing other
    self-similar structures~\cite{wikimediacommons2012subdivision} within reach.

    \item Silica bi-layers~\cite{silica_bilayers}, glass~\cite{sructure_of_2d_glass}, and
    materials that behave like assemblies of 2D particles under
    non-overlap constraints, i.e, like jammed disks on the plane~\cite{jammed_disks}. See Section \ref{sec:bodypin}.
\end{enumerate}
%
In order to study structural and mechanical properties
of a material layer, it is natural to model a material layer as a solution or
realization of a geometric constraint system of appropriate types of geometric
primitives, under metric or algebraic constraints. Such 2D
\dfn{qusecs} (quasi-uniform or self-similar constraint systems, defined
formally in Section \ref{sec:prelim}) can be
used to understand or design material layers (their solutions) with
desired properties.
%
\subsection{Previous Work on Relevant 2D Geometric Constraint Systems}
We now briefly survey existing techniques for studying 2D
qusecs,  many of which are \dfn{bar-joint} systems
(Examples 1, 2 above, see Sections \ref{sec:prelim}, \ref{sec:DRP}, \ref{sec:recomb}),
\dfn{body-hyperpin} systems (Example 4,5, see Section
\ref{sec:bodypin}) or \dfn{pinned-line incidence} systems (Example
3, see Section \ref{sec:pinnedline}). The limitations of these
techniques directly motivate the contributions of this paper.

@@

The contributions of this paper are the following.
\begin{itemize}
  \item We navigate the NP-hardness barrier mentioned in (2) above,
  for finding optimal DR-plans by defining a so-called \dfn{canonical}
  DR-plan and showing a strong Church-Rosser property: \vemph{all
  canonical DR-plans for isostatic or under-constrained 2D qusecs are
  optimal} (Section \ref{sec:DRP}).

  \item We give an efficient ($O(n^2)$) algorithm to find a canonical
  (and hence optimal) DR-plan for all 3 types of 2D qusecs mentioned
  above (Sections \ref{sec:DRP}, \ref{sec:bodypin}, and
  \ref{sec:pinnedline}). The canonical DR-plan elucidates the essence
  of the NP-hardness of finding optimal DR-plans for over-constrained
  systems.

  \item We give a method to deal with the algebraic complexity of
  recombining the realizations or solutions of child subsystems into a
  solution of the parent system
\cite{sitharam2010optimized,sitharam2006well,sitharam2010reconciling}
  Specifically, we define the problem
  of minimally modifying the indecomposable recombination system so
  that it becomes decomposable via a small DR-plan and yet preserves
  the original solutions in an efficiently searchable manner. We show
  formal connection to well known problems such as optimal completion
  of under-constrained systems
\cite{joan-arinyo2003transforming,sitharam2005combinatorial,gao2006ctree}
  and to find paths within the connected components. When the
  modifications are bounded, we obtain new, efficient algorithms for
  realizing both isostatic and under-constrained qusecs by leveraging
  recent results about Cayley parameters in
\cite{sitharam2010convex,sitharam2011cayleyI,sitharam2011cayleyII}
  (Section
  \ref{sec:recomb}).

  \item We briefly describe applications of the above techniques to
  modeling, analyzing, and designing specific properties in 2D material
  layers. For Examples 1,2 (achieving isostaticity, distribution of
  stresses in self-similar, bar-joint systems); Example 3 (canonical
  and optimal DR-plans for pinned line incidence systems
  \cite{sitharam2014incidence})
  and Examples 4,5 (boundary-conditions for achieving various desired
  properties of body-hyperpin systems) \uncited.

  \item We intend to make software implementation and videos available
  upon request and publicly available for the final version of the
  paper.
  \todo{Provide Link}.
\end{itemize}

>>>>

Examples of such materials (See Figure \ref{fig:material_examples})
include:
%
\begin{enumerate}
    \item \label{materialexample1} Cross-sections of microtubule structures~\cite{microtubule_necklace} (Figure \ref{fig:material_examples:microtubule}) e.g., in ciliary membranes and transitions~\cite{microtubule_cilia}.

    \item \label{materialexample2} Cross-sections of organic tissue with hierarchical structure, e.g. compact bone and tendon (Figure \ref{fig:material_examples:tendon}).

    \item \label{materialexample3} Crosslinked cellulose or collagen microfibril monolayers e.g., in cell-walls~\cite{wikimediacommons2010afm}~\cite{wikimediacommons2007plant}, as well as crosslinked actin filaments in the cytoskeleton matrix. See Section \ref{sec:pinnedline}.

    \item \label{materialexample4} More recent, engineered examples, including disordered
    graphene layers~\cite{Graphene1}~\cite{Graphene2} sometimes
    reinforced by microfibrils; and DNA assemblies including a recent
    Sierpinski gasket~\cite{self_assembly_sierpinski}, bringing other
    self-similar structures~\cite{wikimediacommons2012subdivision} within reach.

    \item \label{materialexample5} Silica bi-layers~\cite{silica_bilayers}, glass~\cite{sructure_of_2d_glass}, and
    materials that behave like assemblies of 2D particles under
    non-overlap constraints, i.e, like jammed disks on the plane~\cite{jammed_disks}. See Section \ref{sec:bodypin}.
\end{enumerate}
%
In order to study structural and mechanical properties
of a material layer, it is natural to model a material layer as a solution or
realization of a geometric constraint system of appropriate types of geometric
primitives, under metric or algebraic constraints. Such 2D
\dfn{qusecs} (quasi-uniform or self-similar constraint systems, defined
formally in Section \ref{sec:prelim}) can be
used to understand or design material layers (their solutions) with
desired properties.
%
\subsection{Previous Work on Relevant 2D Geometric Constraint Systems}
We now briefly survey existing techniques for studying 2D
qusecs,  many of which are \dfn{bar-joint} systems
(Examples 1, 2 above, see Sections \ref{sec:prelim}, \ref{sec:DRP}, \ref{sec:recomb}),
\dfn{body-hyperpin} systems (Example 4,5, see Section
\ref{sec:bodypin}) or \dfn{pinned-line incidence} systems (Example
3, see Section \ref{sec:pinnedline}). The limitations of these
techniques directly motivate the contributions of this paper.

@@

The contributions of this paper are the following.
\begin{itemize}
  \item In Section \ref{sec:DRP}, we navigate the NP-hardness barrier mentioned in (2) above,
  for finding optimal DR-plans by defining a so-called \dfn{canonical}
  DR-plan and showing a strong Church-Rosser property: \vemph{all
  canonical DR-plans for isostatic or under-constrained 2D qusecs are
  optimal}.

  \item Also in Section \ref{sec:DRP}, we give an efficient (\candrpcomplexity) algorithm to find a canonical
  (and hence optimal) DR-plan for all 3 types of 2D qusecs mentioned
  above (Sections \ref{sec:DRP}, \ref{sec:bodypin}, and
  \ref{sec:pinnedline}). The canonical DR-plan elucidates the essence
  of the NP-hardness of finding optimal DR-plans for over-constrained
  systems. Futhermore, our optimal/canonical DR-plan satisfies desirable properties such as Cluster Minimality previously studied \cite{hoffman2001decompositionI}.

  \item In Section \ref{sec:recomb}, we give a method to deal with the
  algebraic complexity of recombining the realizations or solutions of
  child subsystems into a solution of the parent system
  \cite{sitharam2010optimized,sitharam2006well,sitharam2010reconciling}
  Specifically, we define the problem of minimally modifying the
  indecomposable recombination system so that it becomes decomposable
  via a small DR-plan and yet preserves the original solutions in an
  efficiently searchable manner. In Section \ref{sec:table}, we show formal connection to well
  known problems such as optimal completion of under-constrained
  systems
  \cite{joan-arinyo2003transforming,sitharam2005combinatorial,gao2006ctree}
  and to find paths within the connected components. When the
  modifications are bounded, we obtain new, efficient algorithms for
  realizing both isostatic and under-constrained qusecs by leveraging
  recent results about Cayley parameters in
  \cite{sitharam2010convex,sitharam2011cayleyI,sitharam2011cayleyII} (see Sections \ref{sec:2-tree-reduction} and \ref{sec:tdecomp}).

  \item In Section \ref{sec:bodypin} and \ref{sec:pinnedline}, we
  briefly describe applications of the above techniques to modeling,
  analyzing, and designing specific properties in 2D material
  layers~\cite{Jackson2008bodypin}. For Examples
  \ref{materialexample1}, \ref{materialexample2} (achieving
  isostaticity, distribution of stresses in self-similar, bar-joint
  systems); Example \ref{materialexample3} (canonical and optimal DR-
  plans for pinned line incidence systems
  \cite{sitharam2014incidence}) and Examples \ref{materialexample4},
  \ref{materialexample5} (boundary-conditions for achieving various
  desired properties of body-hyperpin systems).

  \item We intend to make software implementation and videos available
  upon request and publicly available for the final version of the
  paper.
\end{itemize}

\noindent
\note All proofs appear in Appendix A (\ref{sec:appendix:a}).

### 2014-12-20 doc/GMP2015/s03_drp.tex
Changed proof of algo. Corrected references. Increased references in intro. Changed complexity of algo throughout paper. Added citation in intro. Other smaller changes.

\begin{theorem}\label{theorem:algo_complexity}
    There exists an $O(|V|^3)$ algorithm to find a canonical DR-plan for any 2D isostatic bar-joint graph.
\end{theorem}

\begin{remark}
    while the canonical DR-plan is optimal only if the input graph is independent, if it has only $k$ overconstraints for some fixed $k$, we can still find the optimal DR-plan. However, the time complexity is exponential in $k$.
\end{remark}

>>>>

\begin{theorem}\label{theorem:algo_complexity}
    There exists an \candrpcomplexityv algorithm to find a canonical DR-plan for any 2D isostatic bar-joint graph.
\end{theorem}

\begin{remark}
    While the canonical DR-plan is optimal only if the input graph is independent, if it has only $k$ overconstraints for some fixed $k$, we can still find the optimal DR-plan. However, the time complexity is exponential in $k$.
\end{remark}
